---
title: "Explorative Analysis"
author: "Ayyaz Azeem"
date: "11/15/2020"
output:
  pdf_document: default
  html_document: default
subtitle: Discovering patterns by visualising distriubtions and relationships
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Task 01

```{r , include=FALSE}

#Loading the libaries

library(ggplot2)
library(ggthemes)
library(nlme)
library(gganimate)
library(gapminder)
library(ggExtra)
library(psych)
library(reshape2)
library(dplyr)
library(nycflights13)
library(ggcorrplot)
library(waffle)
library(tidyr)
library(scales)
library(ggalt)
library(data.table)
library(extrafont)
library(lubridate)
library(DT)
library(grid)
library(gridExtra)
library(prettydoc)
#library(devtools)
#library(tidyverse)
library(ggdark)
#library(here)
#library(png)
library(gifski)
library(forcats)
library(tufte)
library(colorspace)
library(viridisLite)
#library(Zelig)
library(formatR)
library(DiagrammeR)
library(xaringan)
library(ggridges)
library(GGally)

```
Exercise 01
Now its time to create your own theme. You are free to chose one theme as for example the dark theme and make changes on top.
Think about your colors, the style of the axis labels, lines and ticks and the legend.

```{r , include=FALSE}

#Defining the theme
 
fill_color = '#111111'
decoration_color = '#cccccc'
main1_color = '#f20675'
main2_color = '#1ce3cd'
main3_color = '#03A062'

theme_set(dark_theme_gray()+ theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title = element_blank(),
  axis.text = element_text(colour = decoration_color, size = 9),
  axis.ticks = element_blank(),
  legend.title = element_blank(),
  panel.background =element_rect(fill = fill_color),
  strip.background =element_rect(fill = fill_color), 
  plot.background = element_rect(fill = fill_color),
  legend.background = element_rect(fill = fill_color)
))

```

### Exercise 02
Choose a different variable like for example population or GDP per capita and check in the same way its frequency using a frequency plot.

```{r}
# variable = gdp Per Capita
#Color set - color brewer
#Adding color as a visual encoding
ggplot(gapminder, aes(gdpPercap, colour = continent)) +
  geom_freqpoly(binwidth = 5000) +
  scale_color_manual(values=c("#478adb", "#cccccc", "#f20675", "#bcc048", "#1ce3cd")) 
 

```

### Exercise 03
Choose a different variable like for example population or GDP per capita and check in the same way its frequency now using a histogram.

```{r}
#ggplot(gapminder, aes(gdpPercap)) + 
#  geom_histogram(colour = "black", fill = main3_color, binwidth = 5000)

ggplot(gapminder, aes(gdpPercap, color = continent)) +
  geom_histogram(colour=decoration_color, fill = decoration_color, alpha = 0.2, size =0) +
  geom_freqpoly(binwidth=5000)+ 
  scale_colour_manual(values=c("#478adb", "#f20675", "#1ce3cd", "#bcc048", "#cccccc"))

```

### Exercise 04
Choose a different variable like for example population or GDP per capita and check in the same way its frequency now using a boxplot.

```{r}

# Boxplot using to numeric variable, we need to define a grouping rule
ggplot(gapminder, aes( gdpPercap, lifeExp)) +
  geom_boxplot(aes(group = cut_width(gdpPercap, 5000)), color=main2_color)

```

### Exercise 05
Choose a different variable like for example population or GDP per capita and check in the same way its frequency now using a density plot.

```{r}
#Small multiple density for carat by the different cuts 
ggplot(gapminder, aes(gdpPercap, stat(density), fill=continent)) + 
  geom_density(color = NA) +
  scale_fill_manual(values=c("#478adb", "#cccccc", "#f20675", "#bcc048", "#1ce3cd")) +   
  facet_wrap(. ~ continent)+
  xlim(-500, 40000)
```

### Exercise 06
Chose two variables from he gapminder dataset and check their relationship by using a scatter plot.

```{r}
#Axis labeling depending on the quantiles for logaritmic scaling
ggplot(gapminder, aes(x=lifeExp, y=gdpPercap)) +
  geom_point(size=0.02, alpha=0.09, color=main3_color)  + 
  xlab("")+ 
  ylab("")+ 
  scale_x_log10(breaks = round(as.vector(quantile(diamonds$carat)), digits = 1))+
  scale_y_log10(breaks = round(as.vector(quantile(diamonds$price)), digits = 1))

```

### Exercise 07
Chose two variables from the gapminder dataset and create in the same way the marginal scatter plots.

```{r}
pp <- ggplot(gapminder, aes(x=lifeExp, y=gdpPercap)) +
  geom_point(color=main3_color) + 
  theme(axis.title=element_blank(), axis.text=element_blank())

ggMarginal(pp, type = "density", fill = main3_color, alpha=1, color='transparent')

```

### Exercise 08
Chose the variables continent and any numeric variable from the gapminder dataset and create the beeswamp plots.
```{r}
library(ggbeeswarm)
ggplot(gapminder, aes(x=gdpPercap, y=continent, colour=continent)) + 
  geom_quasirandom(alpha=0.7, groupOnX=FALSE) +
  scale_colour_manual(values=c("#cccccc", "#478adb", "#f20675", "#1ce3cd", "#bcc048"))

```

### Exercise 09
Chose two variables from the gapminder dataset and check their relationship by using the hexagonal binning.

```{r}
#Aggregation through hexagonal binning - logarithmic scaling 
ggplot(gapminder, aes(x=lifeExp, y=gdpPercap)) +
  geom_hex(alpha = 0.6) +
  scale_x_log10(breaks = round(as.vector(quantile(gapminder$lifeExp)), digits = 1))+
  scale_y_log10(breaks = round(as.vector(quantile(gapminder$gdpPercap)), digits = 1))+
  scale_fill_gradient(low=main3_color, high=main1_color) 

```

### Exercise 10
Use the examples and play around a bit with the options in order to get familiar with scales and axes.

```{r}
# 1
#ggplot(gapminder, aes(x=gdpPercap, y=continent, colour=continent)) + 
#  geom_quasirandom(alpha=0.7, groupOnX=FALSE, method = "smiley") +
#  scale_colour_manual(values=c("#cccccc", "#478adb", "#f20675", "#1ce3cd", #"#bcc048"))+
#  xlim(-500, 40000)
# 2
#ggplot(gapminder, aes(x=lifeExp, y=gdpPercap)) +
#  geom_point(color=main3_color)+
#  xlim(30,80)+
#  ylim(-100,40000)
#Apply another scale to color  
ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop,  color = continent)) +
  geom_point() +
  scale_x_log10()  +
  scale_size(range = c(2, 12)) +
  scale_colour_manual(values = continent_colors) +
  scale_colour_manual(values=c("#478adb", "#f20675", "#1ce3cd", "#bcc048", "#cccccc"))

```

### Exercise 11
Now chose a different variable from the gapminder dataset and do the same analysis flow, feel free to check on another continent as well.

```{r}
# Select only Europe in order to understand which country is the outlier
asia <- dplyr::filter(gapminder, continent == "Asia")

ggplot(asia, aes(year, gdpPercap)) +
  geom_line(color=main3_color) +
  facet_wrap(~country) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Changes in Life Expectancy by country in asia") 

#We can also show the trend as dots 
ggplot(asia, aes(year, gdpPercap)) +
  geom_point(color="grey", size=2) +
  facet_wrap(~country) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Changes in Life Expectancy by country in asia") 

#Coming back to the general checking on patterns globally 

#What will be the output of this code?
trend_color='#FF0000'
#Adding a trend line - defining the method as loss 
ggplot() +
  geom_line(data=gapminder, aes (year, gdpPercap, group = country), lwd = 0.3, show.legend = FALSE, color= decoration_color) + 
  facet_wrap(~ continent, ncol=5, strip.position = "bottom") +
  geom_smooth(data=gapminder, aes(year, gdpPercap, group = 1), lwd = 1.5, method = 'loess', span = 2, se = TRUE, color = trend_color) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Life expectancy by continent including trendline")

#We can even add all data in the background by setting the variable we do the facet with to zero 
ggplot() +
  geom_line(data = transform(gapminder, continent = NULL), aes (year, gdpPercap, group = country), alpha = 0.5, lwd = 0.1, colour = "white") +
  geom_line(data=gapminder, aes (year, gdpPercap, group = country), lwd = 0.3, show.legend = FALSE, color= main3_color) +
  geom_smooth(data=gapminder, aes(year, gdpPercap, group = 1), lwd = 1.5, method = 'loess', span = 0.1, se = TRUE, color = trend_color) + 
  facet_wrap(~ continent, ncol=5, strip.position = "bottom") + 
  theme(strip.background = element_blank(), strip.placement = "outside") +
  theme(axis.text.x = element_blank()) +
  labs(title = "Life expectancy by continent including trendline, showing all data in the back") 

#Now we could filter again on Europe and have far more context 
ggplot() +
  geom_line(data = transform(gapminder, continent = NULL), aes (year, gdpPercap, group = country), alpha = 0.5, lwd = 0.1, colour = "white") +
  geom_line(data=asia, aes (year, gdpPercap, group = country), lwd = 0.3, show.legend = FALSE, color= main3_color) +
  geom_smooth(data=asia, aes(year, gdpPercap, group = 1), lwd = 1.5, method = 'loess', span = 0.1, se = TRUE, color = trend_color) +
  theme(strip.background = element_blank(), strip.placement = "outside") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Life expectancy by country in asia, including a trend line and showing all data in the back") 


```

### Exercise 12
Create different versions of parallel coordinates charts using the gapminder dataset. Also try small multiple.

```{r}
#Simple chart
ggparcoord(gapminder, columns = 4:6, alphaLines = 0.3)

#Simple chart, adding a color code 
ggparcoord(gapminder, columns = 4:6, groupColumn = 2, alphaLines = 0.3) + 
  scale_color_manual(values=c("#478adb", "#f20675", "#1ce3cd", "#bcc048", "#cccccc"))

#Simple chart, adding a color code 
ggparcoord(gapminder, columns = 4:6, groupColumn = 2, alphaLines = 0.3, boxplot = TRUE) + 
  scale_color_manual(values=c("#478adb", "#f20675", "#1ce3cd", "#bcc048", "#cccccc"))

#Showing points, changing transparency and color
ggparcoord(gapminder, columns = 4:6, groupColumn = 2, order = "anyClass",
           showPoints = TRUE, alphaLines = 0.3)+ 
  scale_color_manual(values=c("#478adb", "#f20675", "#1ce3cd", "#bcc048", "#cccccc"))

#Showing points, changing transparency and color
ggparcoord(gapminder, columns = 4:6, groupColumn = 2, order = "anyClass",
           showPoints = TRUE, alphaLines = 0.3)+ 
  scale_color_manual(values=c("#478adb", "#f20675", "#1ce3cd", "#bcc048", "#cccccc"))

#Different way of scaling: No scaling
ggparcoord(gapminder, columns = 4:6, groupColumn = 2, order = "anyClass",
           showPoints = TRUE, alphaLines = 0.3, scale="globalminmax") + 
  scale_color_manual(values=c("#478adb", "#f20675", "#1ce3cd", "#bcc048", "#cccccc"))

#Different way of scaling: Standardize to Min = 0 and Max = 1
ggparcoord(gapminder, columns = 4:6, groupColumn = 2, order = "anyClass",
           alphaLines = 0.3, scale="uniminmax") + 
  scale_color_manual(values=c("#478adb", "#f20675", "#1ce3cd", "#bcc048", "#cccccc"))+   
  facet_wrap(. ~ continent) 

#Different way of scaling: Normalize univariately (substract mean & divide by sd)
ggparcoord(gapminder, columns = 4:6, groupColumn = 2, order = "anyClass",
           showPoints = TRUE, alphaLines = 0.3, scale="std") + 
  scale_color_manual(values=c("#478adb", "#f20675", "#1ce3cd", "#bcc048", "#cccccc"))

#Different way of scaling: Standardize and center variables
ggparcoord(gapminder, columns = 4:6, groupColumn = 2, order = "anyClass",
           showPoints = TRUE, alphaLines = 0.3, scale="center")+ 
  scale_color_manual(values=c("#478adb", "#f20675", "#1ce3cd", "#bcc048", "#cccccc"))
```

### Exercise 13
Now choose a different continent than Europe and also sort by 2007 instead of 1952.

```{r}
#Creating a subsample 
years <- filter(gapminder, year %in% c(1952, 2007)) %>% select(country, continent, year, gdpPercap)

#Check the data 
names(years)
head(years, n=10)
str(years)
summary(years)

#Convert data to wide format
years2 <- spread(years, year, gdpPercap)
names(years2) <- c("country", "continent", "y1952", "y2007")

years3 <- arrange(years2, desc(y2007))
years3$country <- factor(years3$country, levels=rev(years3$country))
#Create a simple dumbbell plot
ggplot(years3, aes(country, x=y1952, xend=y2007))+
  geom_dumbbell(colour=main2_color, size_x=0, size_xend=0)

#Create a simple dumbbell plot
ggplot(years3, aes(country, x=y1952, xend=y2007, color=continent))+
  scale_color_manual(values=c("#478adb", "#cccccc", "#f20675", "#bcc048", "#1ce3cd"))+
  geom_dumbbell(size_x=0, size_xend=0, dot_guide=FALSE, dot_guide_size=0.2, dot_guide_colour=decoration_color)


#Create a simple dumbbell plot
ggplot(years3, aes(country, x=y1952, xend=y2007, color=continent))+
  scale_color_manual(values=c("#478adb", "#cccccc", "#f20675", "#bcc048", "#1ce3cd"))+
  geom_dumbbell(size_x=0, size_xend=0, dot_guide=FALSE, dot_guide_size=0.2, dot_guide_colour=decoration_color) +
  facet_wrap(. ~ continent, ncol=5) 
```

### Exercise 14
Create your own small multiple waffle chart by using different data and a nwe style.

```{r}
iron(
  
  waffle(
    c(men = 62, woman = 38), rows = 5, size = 0.2,
    colors = c(decoration_color, main2_color),
    keep = FALSE,
    title = "WOMAN Arcada 2018", 
    legend='none'),
  
  waffle(
    c(men = 47, woman = 53), rows = 5, size = 0.2,
    colors = c(decoration_color, main2_color),
    keep = FALSE,
    title = "WOMAN Arcada 2019", 
    legend='none'),
  
  waffle(
    c(men = 36, woman = 64), rows = 5, size = 0.2,
    colors = c(decoration_color, main2_color),
    keep = FALSE,
    title = "WOMAN Arcada 2020",
    legend_pos = "bottom")
  
)
```

### Exercise 15
Put all the code of the results of the exercise into a markdown file ansd save it as a pdf.

```{r}
```

## Task2

```{r }
adults<-read.csv("D:/coding/R/DP_LIVE_16112020170746305.csv")

names(adults)
head(adults, n=10)
str(adults)
summary(adults)

#1 General plot definition and theme saved in a variable
f <-ggplot(adults, aes(TIME, Value, group=TIME))

h2 <- ggplot(adults, aes(ï..LOCATION, Value, group=INDICATOR))
h2 + geom_line()+theme(axis.text.x = element_text(angle = 45, hjust = 1))

h2 + geom_line() + 
  geom_smooth(aes(), colour = trend_color, size = 1.5, method = "lm", se = FALSE)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

h2 + geom_line() +
  geom_smooth(aes(group = 1), colour = trend_color, size = 3, method = "lm", se = TRUE)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Now we combine a box-plot with the line chart 
h2 <- ggplot(adults, aes(ï..LOCATION, Value))
h2 + geom_boxplot() + geom_line()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# We can add the a line chart again for all the subjects  
h2 + geom_boxplot() + geom_line(aes(group = INDICATOR), size=0.3, colour=trend_color)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# We can add the a line chart grouped 
h2 + geom_boxplot() + geom_smooth(aes(group = 1), method = "lm", se = FALSE, colour=trend_color)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```